---
title: "Download data"
author: "EE Holmes"
output: html_document
---

# Download data from an ERDDAP server

This will download data from the CoastWatch ERDDAP server. Alter the URL to download from a different ERDDAP server or data set.

## Use rerddap

```
require(devtools)
devtools::install_github("ropensci/rerddap")
devtools::install_github("rmendels/rerddapXtracto") 
```

Specify where to save the downloaded data. Data from a `griddap()` call can be verbose, so you'll probably want a folder for it.
```{r create_dir}
fil_dir <- file.path(here::here(), "inst", "extdata")
if(!dir.exists(fil_dir)) dir.create(fil_dir)
```

We will download SST data from CoastWatch.  [Here](https://coastwatch.pfeg.noaa.gov/erddap/griddap/ncdcOisst21Agg_LonPM180.html) is the data access page for that dataset.
```{r}
lats <- c(40.375, 50.375)
lons <- c(-141.875, -120.875)
df_info <- rerddap::info("ncdcOisst21Agg_LonPM180")
df <- rerddap::griddap("ncdcOisst21Agg_LonPM180", latitude = lats, longitude = lons, time = c("2021-06-19", "2021-06-19"), fields = "sst")$data
```

`df` is a data frame with lat and lon.
```{r}
head(df)
```

There is [no projection information](https://coastwatch.pfeg.noaa.gov/erddap/info/ncdcOisst21Agg_LonPM180/index.html) for these data, but the meta data just says that it is on a uniform lat-lon grid. That tells us that at minimum it is `"+proj=longlat"`. It is important to include this information so that we can combine this with other map data later.

## Turn this into a raster

We can then turn this matrix into a raster using the **raster** package and the `rasterFromXYZ()` function. 
```{r}
df2 <- data.frame(x=df$lon, y=df$lat, z=df$sst)
ras <- raster::rasterFromXYZ(df2, crs = "+proj=longlat")
```

## Plot the raster

We can do this with the **raster** package.
```{r}
library(raster)
plot(ras)
```
We can use the **cmocean** package and use its thermal palette.
```{r}
plot(ras, col = cmocean::cmocean("thermal")(100))
```

We can plot with **ggplot2** also.
```{r}
require(ggplot2)
# Plot
gg <- ggplot(df) +
  geom_raster(aes(lon, lat, fill = sst)) +
  scale_fill_gradient2(midpoint = mean(df$sst, na.rm = TRUE),
                       low = "blue",
                       mid = "white",
                       high = "red") +
  labs(x = NULL,
       y = NULL,
       fill = "Celcius",
       title = "Sea Surface Temperature (SST)")
gg
```

It's a bit ugly. We might futz with the look and use the **cmocean** palette.
```{r}
gg + theme_bw() +
  scale_x_continuous(limits = lons, expand = c(-0.01, -0.01)) +
  scale_y_continuous(limits = lats, expand = c(-0.01, -0.01)) +
  cmocean::scale_fill_cmocean(alpha=1) + theme_bw()
```

The default raster plot is a bit deformed since it is long-lat on the x and y axis. We can see what it would look like in a different projection.

```{r}
newcrs <- "+proj=wintri +lon_0=-125 +lat_1=46 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
ras_proj<- projectRaster(ras, crs=newcrs, over=TRUE)
plot(ras_proj)
title(stringr::str_sub(newcrs, 1, 12))
```

## Save

We'll save this to use later. You'll want to change the `data_dir` to wherever you are saving your data. I am saving the data in different forms since in the examples to follow, different forms will be used.
```{r eval=FALSE}
data_dir <- file.path(here::here(), "data", "sample_raster.rda")
sample_raster <- list(df=df, raster=ras, lats=lats, lons=lons)
save(sample_raster, file=data_dir)
```


