---
title: "Add coastline"
author: "EE Holmes"
output: html_document
---

This example assumes that you have a sample raster. You can load the sample one or go to the download data vignette to download your own. `sample_raster` is a demo SST raster and data frame.

Load the sample data.
```{r}
data("sample_raster", package="basics")
df <- sample_raster$df
ras <- sample_raster$raster
lons <- sample_raster$lons
lats <- sample_raster$lats
```

Load the needed packages for plotting.
```{r}
require(raster)
require(ggplot2)
```


## Download a coastline

There are a variety of places you can get a coastline.

You can download via **raster**.
```{r eval=FALSE}
coast <- raster::getData("GADM", country = "USA", level = 1)
wa_or_coast <- subset(usashp, NAME_1 %in% c("Washington", "Oregon"))
```

Or you could get it from **rnaturalearth** which is quite a bit faster. With scale=50, the coastline has some detail. You could pass in scale of 110 or 10.
```{r}
coast <- rnaturalearth::ne_coastline(scale = 50, returnclass = "sp")
```

## Crop and plot

I'll use **rnaturalearth**. The coast just downloaded is for the whole world. We'll want to crop that down to our region. Note I need to use `library(raster)` so that I have access to the plot methods for spatial objects.
```{r}
library(raster)
wa_or_coast <- raster::crop(coast, raster::extent(lons[1], lons[2], lats[1], lats[2]))
plot(wa_or_coast)
```


## Add on raster with base R

```{r}
plot(ras)
plot(wa_or_coast, add=TRUE)
```

We can plot with **ggplot2** also.
```{r}
require(ggplot2)
# Plot
gg <- ggplot(df) +
  geom_raster(aes(lon, lat, fill = sst)) +
  scale_fill_gradient2(midpoint = mean(df$sst, na.rm = TRUE),
                       low = "blue",
                       mid = "white",
                       high = "red") +
  labs(x = NULL,
       y = NULL,
       fill = "Celcius",
       title = "Sea Surface Temperature (SST)")
```

The way that **ggplot2** works is to run `fortify()` on the SpatialLines object to create a data frame. Then we use `geom_path()` to plot that. But if you look at the coast, you see lots of islands. We need to tell `geom_path()` that there are these groups of paths in the data frame.
```{r}
gg <- gg + 
  geom_path(data=wa_or_coast,  aes(x=long,y=lat, grouping=id), size=1, na.rm=TRUE)
gg
```

## **cmocean** palette

Let's use the **cmocean** package to use it's thermal palette.

```{r}
library(cmocean)
gg + scale_fill_cmocean(alpha=1) + theme_bw()
```

We can use this for our raster plot too.
```{r}
plot(ras, col = cmocean("thermal")(100))
```


## Add a line parallel to the coastline

```{r}
coast110 <- rnaturalearth::ne_coastline(scale = 110, returnclass = "sp")
coast110 <- raster::crop(coast110, raster::extent(lons[1], lons[2], lats[1], lats[2]))
offcoast <- raster::shift(coast110, dx=-2)
```

```{r}
plot(ras)
plot(wa_or_coast, add=TRUE)
plot(offcoast, add=TRUE, col="red")
```


The default raster plot is a bit deformed since it is long-lat on the x and y axis. We can see what it would look like in a different projection.
```{r}
newcrs <- "+proj=wintri +lon_0=-125 +lat_1=46 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
ras_win <- projectRaster(ras, crs=newcrs, over=T)
plot(ras_win)
plot(spTransform(wa_or_coast, newcrs), add=TRUE)
title(stringr::str_sub(newcrs, 1, 10))
```
