---
title: "3 - Stamen Map"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3 - Stamen Map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example assumes that you have a sample raster. You can load the sample one or go to the download data vignette to download your own. `sample_raster` is a demo SST raster and data frame.

Load the sample data.
```{r}
data("sample_raster", package="basics")
df <- sample_raster$df
ras <- sample_raster$raster
lons <- sample_raster$lons
lats <- sample_raster$lats
```

Load the needed packages for plotting.
```{r}
require(raster)
require(ggplot2)
```


## Stamen map the gridded data

```{r}
# Give coordinates for map
bbox <- c(left = lons[1], bottom = lats[1], right = lons[2], top = lats[2])
# Get stamen map
ocean_map <- ggmap::get_stamenmap(bbox, zoom = 5, maptype = "terrain")
```

If you wanted to change the color of the ocean, you could do this.
```{r eval=FALSE}
attr_om <- attr(ocean_map, "bb")    # save attributes from original

# change ocean color in raster
ocean_map[ocean_map == "#99B3CC"] <- "#F5F5F5"
class(ocean_map) <- c("ggmap", "raster")
attr(ocean_map, "bb") <- attr_om
```

```{r}
require(ggmap)
# Plot
gg <- ggmap(ocean_map) +
  geom_tile(aes(lon, lat, fill = sst),
            data = df, na.rm = TRUE) +
  scale_fill_gradient2(midpoint = mean(df$sst, na.rm = TRUE),
                       low = "blue",
                       mid = "white",
                       high = "red",
                       na.value = "transparent") +
  labs(x = NULL,
       y = NULL,
       fill = "Celcius",
       title = "Sea Surface Temperature (SST)")
gg
```


## Adding a coastline

Download a coast
```{r}
coast <- rnaturalearth::ne_coastline(scale = 50, returnclass = "sp")
wa_or_coast <- raster::crop(coast, raster::extent(lons[1], lons[2], lats[1], lats[2]))
plot(wa_or_coast)
```

The way that **ggplot2** works is to run `fortify()` on the SpatialLines object to create a data frame. Then we use `geom_path()` to plot that. But if you look at the coast, you see lots of islands. We need to tell `geom_path()` that there are these groups of paths in the data frame.
```{r}
library(ggmap)
gg <- gg + geom_path(data=wa_or_coast,  aes(x=long,y=lat, grouping=id), size=1, na.rm=TRUE)
gg
```

We might want to futz with the look a bit
```{r}
gg +
  scale_x_continuous(limits = lons, expand = c(-0.01, -0.01)) +
  scale_y_continuous(limits = lats, expand = c(-0.01, -0.01)) +
  theme_bw()
```


