---
title: "Add lines parallel to coast"
author: "EE Holmes"
output: html_document
---

This example shows how to make a line that is parallel to the coast and how to make a line that follows the coast but removes any inlets (like Puget Sound or San Francisco Bay).

Load the sample data.
```{r}
data("sample_raster", package="basics")
df <- sample_raster$df
ras <- sample_raster$raster
lons <- sample_raster$lons
lats <- sample_raster$lats
```

Load the needed packages for plotting.
```{r}
require(raster)
require(ggplot2)
```

```{r echo=FALSE}
only.holes <- function (x) 
{
  if (!any(which(utils::installed.packages()[, 1] %in% "maptools"))) 
    stop("please install maptools package before running this function")
  xp <- slot(x, "polygons")
  holes <- lapply(xp, function(x) sapply(methods::slot(x, "Polygons"), 
                                         methods::slot, "hole"))
  res <- lapply(1:length(xp), function(i) methods::slot(xp[[i]], 
                                                        "Polygons")[holes[[i]]])
  IDs <- row.names(x)
  x.fill <- sp::SpatialPolygons(lapply(1:length(res), function(i) sp::Polygons(res[[i]], 
                                                                               ID = IDs[i])), proj4string = sp::CRS(sp::proj4string(x)))
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), maptools::checkPolygonsHoles)
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), "comment<-", NULL)
  pids <- sapply(methods::slot(x.fill, "polygons"), function(x) methods::slot(x, "ID"))
  x.fill <- sp::SpatialPolygonsDataFrame(x.fill, data.frame(row.names = pids, 
                                                            ID = 1:length(pids)))
  return(x.fill)
}
```

## Download the world coastlines

For this buffer, I am using a fairly zoomed out version of the coastline. You could use the coastline from **rnaturalearth**.

## Create a line around the world's coastlines at 300 km offshore

This is a projection in meters (not longlat). Note that I have set the units to km. See comments in the **add_coast_buffer** vignette about why `rgeos::gBuffer()` must be used and not `raster::buffer()`.

* Select a project in meters not longlat
* Transform our world polygon in that projection
* Create a buffer polygon around the world polygon
* Erase the interior world polygon (not sure one has to do that but you do)
* Clean up by removing some of the small holes in the buffer polygon

```{r}
# Step 1
newcrs <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
# Step 2 World coastline in meters
data("world", package="basics")
mworld <- sp::spTransform(world, newcrs)
# Step 3
buff300 <- rgeos::gBuffer(mworld, width = 300, byid=TRUE)
# Step 4
e <-raster::erase(buff300, mworld)
```

We want just the outer part of this polygon. We will use the `remove.holes()` function from **spatialEco**.
```{r}
# Step 5
e300 <- spatialEco::remove.holes(spatialEco::remove.holes(e))
```

Now we can plot and the holes that were removed are the red lines.
```{r}
plot(e, border="red")
plot(e300, add=TRUE)
plot(mworld, add=TRUE, col="grey")
```

```{r}
pts <- sp::SpatialPoints(cbind(lons, lats), proj4string = sp::CRS("+proj=longlat"))
mpts <- sp::spTransform(pts, newcrs)
plot(e300, border="red", axes=TRUE, xlim=mpts@bbox[1,], ylim=mpts@bbox[2,])
plot(mworld, add=TRUE, col="grey")
```

## Create a line around the world's coastlines at 10 km offshore

We could use the method above but the 20 km buffer would dip into the inlets and we don't want that. We want this to be 280 km inside of the 300km line.

* Turn `e300` into lines. Otherwise our 280 km buffer will only be outside and we want the inner one.
* Create a 290 km buffer to that
* Keep ONLY the holes on that new buffer object.

```{r}
el300 <- as(e300, "SpatialLines")
buff20 <- rgeos::gBuffer(el300, width=280)
```

If we plot this we see that it is on both sides of our 300 km line. We want only the inner lines.
```{r}
plot(buff20)
plot(el300, add=TRUE, col="red")
```


Remove the outer part of the polygons. Let's make it lines too.
```{r}
e20 <- only.holes(buff20)
el20 <- as(e20, "SpatialLines")
```

Now it looks better. Potentially we'll want to remove the smaller polygons (islands).
```{r}
plot(el20)
plot(el300, add=TRUE, col="red")
```

We can look off the WA coast.
```{r}
plot(el20, col="black", axes=TRUE, xlim=mpts@bbox[1,], ylim=mpts@bbox[2,])
plot(el300, add=TRUE, col="red")
plot(mworld, add=TRUE, col="grey")
```

## Add sample locations

Now we can add some locations where we will sample the coastline

```{r}
numOfPoints  <-  rgeos::gLength(el20) / 100
sample.pts <- sp::spsample(el20, n = numOfPoints, type = "regular")
```

```{r}
plot(el20, col="black", axes=TRUE, xlim=mpts@bbox[1,], ylim=mpts@bbox[2,])
plot(el300, add=TRUE, col="red")
plot(mworld, add=TRUE, col="grey")
plot(sample.pts, add=TRUE, pch=1)
title("sample points every 100 km")
```


## Only Holes function direct from `spatialEco::remove.holes()`
```{r}
only.holes <- function (x) 
{
  if (!any(which(utils::installed.packages()[, 1] %in% "maptools"))) 
    stop("please install maptools package before running this function")
  xp <- slot(x, "polygons")
  holes <- lapply(xp, function(x) sapply(methods::slot(x, "Polygons"), 
                                         methods::slot, "hole"))
  res <- lapply(1:length(xp), function(i) methods::slot(xp[[i]], 
                                                        "Polygons")[holes[[i]]])
  IDs <- row.names(x)
  x.fill <- sp::SpatialPolygons(lapply(1:length(res), function(i) sp::Polygons(res[[i]], 
                                                                               ID = IDs[i])), proj4string = sp::CRS(sp::proj4string(x)))
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), maptools::checkPolygonsHoles)
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), "comment<-", NULL)
  pids <- sapply(methods::slot(x.fill, "polygons"), function(x) methods::slot(x, "ID"))
  x.fill <- sp::SpatialPolygonsDataFrame(x.fill, data.frame(row.names = pids, 
                                                            ID = 1:length(pids)))
  return(x.fill)
}
```