---
title: "6 - Find Offshore Nearest Point"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{6 - Find Offshore Nearest Point}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example shows how to select the nearest point that is some distance offshore and then compute some statistics for that point.

Load the sample data.
```{r}
data("sample_raster", package = "basics")
df <- sample_raster$df
ras <- sample_raster$raster
lons <- sample_raster$lons
lats <- sample_raster$lats
```

Load the needed packages for plotting.
```{r}
require(raster)
require(ggplot2)
```

## Preliminaries

### Download the world coastline

```{r}
world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
world <- rgeos::gUnaryUnion(world)
```

### Set up our projection in meters

Create some transformed versions of objects we use a bit.
```{r}
newcrs <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
mworld <- sp::spTransform(world, newcrs)
mras <- raster::projectRaster(ras, crs = newcrs, over = TRUE)
mpts <- sp::SpatialPoints(cbind(lons, lats), proj4string = sp::CRS("+proj=longlat"))
mpts <- sp::spTransform(mpts, newcrs)
```

### Create a line around the world's coastlines at 300 km offshore

See comments in the **add_coast_buffer** vignette about why `rgeos::gBuffer()` must be used and not `raster::buffer()`.
```{r}
buff1 <- rgeos::gBuffer(mworld, width = 300, byid = TRUE)
e <- raster::erase(buff1, mworld)
plot(e)
```

We want just the outer part of this polygon. We will use the `remove.holes()` function from **spatialEco**.
```{r}
e300 <- spatialEco::remove.holes(spatialEco::remove.holes(e))
plot(e300)
```

We can add the world to this and see that it is just the coastline area. We have lost some interior coastlines like Gulf of Mexico.
```{r}
plot(e300)
plot(mworld, add = TRUE, col = "grey")
```

## Find closest point

```{r}
el <- as(mworld, "SpatialLines")
el <- raster::crop(el, mpts@bbox)
set.seed(123)
pts <- sp::spsample(el, n = 1, type = "regular")
```

We want to find the point on the coastline that is closest to this 300m offshore line.
```{r}
plot(e300, border = "red", axes = TRUE, xlim = mpts@bbox[1, ], ylim = mpts@bbox[2, ])
plot(mworld, add = TRUE, col = "grey")
plot(pts, add = TRUE, pch = 19)
```


## Finding the closest point

The `snapPointsToLines()` function in **maptools** should work, but does not seem to handle the types of SpatialLines objects that are returned when we convert our 300 km polygon into lines.

```{r}
el300 <- as(e300, "SpatialLines")
close_pt <- maptools::snapPointsToLines(pts, el300, maxDist = 500)
plot(mworld)
plot(close_pt, add = TRUE, pch = 19, col = "red")
plot(pts, add = TRUE, pch = 19)
```

So we need to use a more brute force method. 

* Convert our 300km polygon to SpatialLines
* Make the SpatialLines object into a dataframe
* Pass that to `maptools::nearestPointOnLine()`
* Make back into SpatialPoints object for plotting

```{r}
el300 <- as(e300, "SpatialLines")
df <- c()
n <- length(el300@lines[[1]]@Lines)
for (i in 1:n) {
  df <- rbind(df, cbind(el300@lines[[1]]@Lines[[i]]@coords, ID = i))
}
close_pt <- maptools::nearestPointOnLine(df, pts@coords)
close_pt <- sp::SpatialPoints(matrix(close_pt, ncol = 2), proj4string = CRS(newcrs))
```

````{r}
plot(e300, border = "red", axes = TRUE, xlim = mpts@bbox[1, ], ylim = mpts@bbox[2, ])
plot(mworld, add = TRUE, col = "grey")
plot(pts, add = TRUE, pch = 19)
plot(close_pt, add = TRUE, pch = 1)
```

## Compute statistics associated with the point

We want to compute some statistics associated with the offshore point.
```{r}
plot(mras, axes = TRUE, xlim = bbox(mras)[1, ], ylim = bbox(mras)[2, ])
plot(e300, border = "red", add = TRUE)
plot(mworld, add = TRUE, col = "grey")
plot(pts, add = TRUE, pch = 19)
plot(close_pt, add = TRUE, pch = 1)
```

Get the SST at that point.
```{r}
raster::extract(mras, close_pt)
```


Get the mean SST in a 100km circle around that point. For this we need to make a circle polygon around the point.
```{r}
circle_pt <- raster::buffer(close_pt, width = 100)
```

```{r echo=FALSE}
plot(mras, axes = TRUE, xlim = bbox(mras)[1, ], ylim = bbox(mras)[2, ])
plot(e300, border = "red", add = TRUE)
plot(mworld, add = TRUE, col = "grey")
plot(pts, add = TRUE, pch = 19)
plot(close_pt, add = TRUE, pch = 19, col = "red")
plot(circle_pt, add = TRUE)
```
We use `raster::extract()` again.
```{r}
vals <- raster::extract(mras, circle_pt)[[1]]
vals
```
and we take the mean.
```{r}
mean(vals, na.rm = TRUE)
```

## Save the objects

```{r eval=FALSE}
crs.wintri <- newcrs
save(crs.wintri, file = file.path(here::here(), "data/crs_wintri.rda"))
world.wintri <- mworld
save(world.wintri, file = file.path(here::here(), "data/world_wintri.rda"))
save(world, file = file.path(here::here(), "data/world.rda"))
buffer300.wintri <- list(line = el300, polygon = e300, df = df, crs = newcrs)
save(buffer300.wintri, file = file.path(here::here(), "data/buffer300_wintri.rda"))
```

## Make a function

```{r}
# Make a buffer and return in various formats
make_buffer <- function(d = 300, units = "km", crs.to.use = "wintri", remove.holes = TRUE) {
  world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
  world <- rgeos::gUnaryUnion(world)
  newcrs <- paste0("+proj=", crs.to.use, " +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=", units, " +no_defs")
  mworld <- sp::spTransform(world, newcrs)
  buff1 <- rgeos::gBuffer(mworld, width = d, byid = TRUE)
  e <- raster::erase(buff1, mworld)
  if (remove.holes) e <- spatialEco::remove.holes(spatialEco::remove.holes(e))
  el <- as(mworld, "SpatialLines")
  df <- c()
  n <- length(el@lines[[1]]@Lines)
  for (i in 1:n) {
    df <- rbind(df, cbind(el@lines[[1]]@Lines[[i]]@coords, ID = i))
  }
  return(list(polygon = e, line = el, df = df, crs = newcrs))
}
```

```{r}
# Find the nearest point to the buffer
get.nearest.buffer.pt <- function(pts, buff = buffer300.wintri$df, newcrs = buffer300.wintri$crs) {
  if (inherits(pts, "SpatialPoints")) pts <- pts@coords
  if (inherits(buff, "SpatialPolygon")) {
    buff <- as(buff, "SpatialLines")
  }
  if (inherits(buff, "SpatialLines")) {
    df <- c()
    n <- length(buff@lines[[1]]@Lines)
    for (i in 1:n) df <- rbind(df, cbind(buff@lines[[1]]@Lines[[i]]@coords, ID = i))
    buff <- df
  }
  close_pt <- maptools::nearestPointOnLine(df, pts)
  close_pt <- sp::SpatialPoints(matrix(close_pt, ncol = 2), proj4string = CRS(newcrs))
  return(close_pt)
}
```

```{r}
# Get the mean raster values around a point
get.mean.around.pt <- function(pts, ras, d = 100, units = "km", newcrs = crs.wintri, fun = "mean") {
  if (!inherits(pts, "SpatialPoints")) stop("pts shoudl be a SpatialPoints object")
  if (!inherits(ras, "raster")) stop("ras should be a raster")
  mpts <- pts
  if (!identical(crs(pts), newcrs)) mpts <- sp::spTransform(mpts, newcrs)
  mras <- ras
  if (!identical(crs(ras), newcrs)) mras <- raster::projectRaster(ras, crs = newcrs, over = TRUE)
  circle_pt <- raster::buffer(pts, width = d)
  vals <- raster::extract(mras, circle_pt)
  val <- c()
  for (i in 1:length(vals)) val <- c(val, do.call(fun, vals[[i]], na.rm = TRUE))
  return(val)
}
```

```{r}
# Get the mean raster values at a point
get.mean.at.pt <- function(pts, ras, newcrs = crs.wintri) {
  if (!inherits(pts, "SpatialPoints")) stop("pts shoudl be a SpatialPoints object")
  if (!inherits(ras, "raster")) stop("ras should be a raster")
  mpts <- pts
  if (!identical(crs(pts), newcrs)) mpts <- sp::spTransform(mpts, newcrs)
  mras <- ras
  if (!identical(crs(ras), newcrs)) mras <- raster::projectRaster(ras, crs = newcrs, over = TRUE)
  val <- raster::extract(mras, pts)
  return(val)
}
```
