---
title: "7 - Compute Stats"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{7 - Compute Stats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example shows how work through the points along the 20km coastline and return the stats for each point.

The **basics** package our coastline, coastal sample points and the offshore 300km line.
```{r}
data(package = "basics")
```

Load the needed packages for plotting and data.
```{r}
require(raster)
require(ggplot2)
require(basics) # the data
```

## Make a function to get the nearest points on the 300 km line

We need to pass it our 300 km line as a data frame (long, lat, ID) along with the long-lat of a point on the coast. Little bumps in the coast and 300km lines cause the nearest points to be clustered around dips toward the coast. Simplifing the line helps prevent this. A `tol` about 50 seems to work. `na.rm` determines if NA are ignored or not.

```{r}
# l is SpatialLines for 300 km line
# p is SpatialPoints of coastal point
NearestPoints <- function(l, p, tol=1) {
  #if(!is.null(tol)) l <- rgeos::gSimplify(l, tol=tol)
  if(tol!=0) l <- smoothr::smooth(l, method="ksmooth", smoothness=5)
  df <- c()
  n <- length(l@lines[[1]]@Lines)
  for (i in 1:n) {
    df <- rbind(df, cbind(l@lines[[1]]@Lines[[i]]@coords, ID = i))
  }
  pts <- p@coords
  close_coords <- matrix(NA, nrow(pts), 2)
  for (i in 1:nrow(pts)) {
    close_coords[i, ] <- maptools::nearestPointOnLine(df, pts[i, ])
  }
  close_sp <- sp::SpatialPoints(close_coords, proj4string = crs(l))
  return(list(coords = close_coords, sp = close_sp))
}
```

## Make a function to get the average SST for these 300 km points

Because the raster might include islands, which would be NA, setting `na.rm=FALSE` means coastal points that are inside of an island will be NA.

```{r}
# p is SpatialPoints the coastal points
# d is the distance (in whatever units p is in)
SST.offshore <- function(r, p, d = 100, na.rm=FALSE) {
  if(d==0){ # Get value at points
  vals <- raster::extract(r, p)
  return(vals)
  }
  # Get values in circle around points
  pts <- p@coords
  vals <- c()
  for(i in 1:nrow(pts)){
    pt <- sp::SpatialPoints(pts[i,,drop=FALSE])
    circle_pt <- raster::buffer(pt, width = d)
    vals <- c(vals, mean(raster::extract(r, circle_pt)[[1]], na.rm=na.rm))
  }
  return(vals)
}
```

## Make a function to get the SST for the coastal points

If `d=0` then it just returns the SST at the point. If `d` is not zero, it takes the average along `l` that is `d` to each side of the points.

```{r}
SST.coast <- function(r, p, l=NULL, d=0) {
  if(d==0){
  vals <- raster::extract(r, p)
  return(vals)
  }
  if(is.null(l)) stop("If d is not zero, need the coastline.")
  pts <- p@coords
  vals <- c()
  for(i in 1:nrow(pts)){
    pt <- sp::SpatialPoints(pts[i,,drop=FALSE])
    circle_pt <- raster::buffer(pt, width = d)
    segment_pt <- raster::intersect(l, circle_pt)
    vals <- c(vals, mean(raster::extract(r, segment_pt)[[1]], na.rm=TRUE))
  }
  return(vals)
}
```

```{r}
upwelling <- function(r, p, l.offshore, l.coast=NULL, d.offshore=100, d.coast=0, threshold=2.5, tol=NULL){
  coast.pts <- p
  offshore.pts <- NearestPoints(l.offshore, coast.pts, tol=tol)$sp
  off.sst <- SST.offshore(r, offshore.pts, d = d.offshore)
  coast.sst <- SST.coast(r, p, l=l.coast, d=d.coast)
  df <- data.frame(p@coords, offshore=off.sst, coast=coast.sst, 
                   upwelling=(off.sst-coast.sst)>=threshold)
  colnames(df) <- c("longitude", "latitude", "offshore.SST", "coast.SST", "upwelling")
  return(list(df=df, offshore.pts=offshore.pts, tol=tol, threshold=threshold, d.offshore=d.offshore, d.coast=d.coast))
}
```


## Now we can apply this to a set of points in our area of interest (WA/OR)

```{r}
coast.pts <- raster::crop(sample_points$km100, 
                       raster::bbox(raster::trim(sample_raster$raster.wintri)))
```

```{r}
out <- upwelling(sample_raster$raster.wintri, coast.pts, 
                 buffer300$wintri$line, d.offshore=50, 
                 d.coast=50, l.coast=buffer20$wintri$line, tol=30)
out$df
```

```{r}
coast.pts <- raster::crop(sample_points$km100, 
                       raster::bbox(raster::trim(sample_raster$raster.wintri)))
out2 <- upwelling(sample_raster$raster.wintri, coast.pts, 
                 raster::crop(buffer300$wintri$line, raster::bbox(sample_raster$raster.wintri)), d.offshore=100, tol=50)
out2$df
```
```{r include=FALSE}
figcap <- paste("Coast points and corresponding offshore points.", 
                ifelse(!is.null(out$tol), paste("tol =", out$tol, "Offshore line is smoothed so points won't fall on line."), "Offshore line is not smoothed."), "The average SST in a circle of radius", out$d.offshore, "was used.")
```
```{r fig.cap=figcap}
plot(sample_raster$raster.wintri)
plot(buffer300$wintri$line, add=TRUE)
plot(buffer20$wintri$line, add=TRUE)
cols <- ifelse(out$df$upwelling, "red", "black")
cols[is.na(cols)] <- "grey"
plot(coast.pts, add=TRUE, pch=19, col=cols)
plot(out$offshore.pts, add=TRUE, pch=1)
title("coast points and corresponding offshore points")
```


