---
title: "Saved Data"
author: "Eli Holmes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Saved Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This example shows how the data in the package were created and saved.

Specify where to save the downloaded data. Data from a `griddap()` call can be verbose, so you'll probably want a folder for it.
```{r create_dir}
fil_dir <- file.path(here::here(), "inst", "extdata")
if(!dir.exists(fil_dir)) dir.create(fil_dir)
```

Winkel Tripel projection
```{r}
crs.wintri <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
```


## `sample_raster`

```{r}
lats <- c(40.375, 50.375)
lons <- c(-141.875, -120.875)
df_info <- rerddap::info("ncdcOisst21Agg_LonPM180")
df <- rerddap::griddap("ncdcOisst21Agg_LonPM180", latitude = lats, longitude = lons, time = c("2021-06-19", "2021-06-19"), fields = "sst")$data
```

We turn this matrix into a raster using the **raster** package and the `rasterFromXYZ()` function. 
```{r}
df2 <- data.frame(x=df$lon, y=df$lat, z=df$sst)
ras <- raster::rasterFromXYZ(df2, crs = "+proj=longlat")
```

Convert to Winkel Tripel projection
```{r}
ras.wintri <- raster::projectRaster(ras, crs=crs.wintri, over=TRUE)
lats.wintri <- raster::bbox(ras.wintri)[1,]
lons.wintri <- raster::bbox(ras.wintri)[2,]
```

```{r}
data_dir <- file.path(here::here(), "data", "sample_raster.rda")
sample_raster <- list(df=df, raster=ras, lats=lats, lons=lons, ras.wintri, lats.wintri, lons.wintri, crs.wintri)
save(sample_raster, file=data_dir)
```

## World coastline

```{r}
world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
# Get rid of interior boundaries
world <- rgeos::gUnaryUnion(world)
```

```{r}
world.wintri <- sp::spTransform(world, crs.wintri)
```

## 300 km offshore line

Note, the units for `world.wintri` were set to km. It is important that this code is run on 
a polygon in a meter as opposed to longlat projection.
```{r}
buff1 <- rgeos::gBuffer(world.wintri, width = 300, byid = TRUE)
# erase the inner world polygon
e <- raster::erase(buff1, world.wintri)
# Use the `remove.holes()` function to get only the outer line.
e300 <- spatialEco::remove.holes(spatialEco::remove.holes(e))
el300 <- as(e300, "SpatialLines")
```

Convert to a dataframe also.
```{r}
df300 <- c()
n <- length(el300@lines[[1]]@Lines)
for (i in 1:n) {
  df300 <- rbind(df300, cbind(el300@lines[[1]]@Lines[[i]]@coords, ID = i))
}
```

Assemble into a list.
```{r}
buffer300.wintri <- list(line = el300, polygon = e300, df = df, crs = crs.wintri)
```

## Make 20km coast line

Only Holes function direct adaptation from `spatialEco::remove.holes()`

```{r}
only.holes <- function (x) 
{
  if (!any(which(utils::installed.packages()[, 1] %in% "maptools"))) 
    stop("please install maptools package before running this function")
  xp <- slot(x, "polygons")
  holes <- lapply(xp, function(x) sapply(methods::slot(x, "Polygons"), 
                                         methods::slot, "hole"))
  res <- lapply(1:length(xp), function(i) methods::slot(xp[[i]], 
                                                        "Polygons")[holes[[i]]])
  IDs <- row.names(x)
  x.fill <- sp::SpatialPolygons(lapply(1:length(res), function(i) sp::Polygons(res[[i]], 
                                                                               ID = IDs[i])), proj4string = sp::CRS(sp::proj4string(x)))
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), maptools::checkPolygonsHoles)
  methods::slot(x.fill, "polygons") <- lapply(methods::slot(x.fill, 
                                                            "polygons"), "comment<-", NULL)
  pids <- sapply(methods::slot(x.fill, "polygons"), function(x) methods::slot(x, "ID"))
  x.fill <- sp::SpatialPolygonsDataFrame(x.fill, data.frame(row.names = pids, 
                                                            ID = 1:length(pids)))
  return(x.fill)
}
```

Create buffer 280 km around the 300km line. Note because we are doing this around a line, we don't need to erase the inner polygon as we had to do for the 300km line.
```{r}
buff20 <- rgeos::gBuffer(el300, width=280)
```

Remove the outer part of the polygons and make into lines.
```{r}
e20 <- only.holes(buff20)
el20 <- as(e20, "SpatialLines")
```

Convert to a dataframe also.
```{r}
df20 <- c()
n <- length(el20@lines[[1]]@Lines)
for (i in 1:n) {
  df20 <- rbind(df20, cbind(el20@lines[[1]]@Lines[[i]]@coords, ID = i))
}
```

## Coast sample locations every 100km

```{r}
numOfPoints  <-  rgeos::gLength(el20) / 100
sample.pts.100km <- sp::spsample(el20, n = numOfPoints, type = "regular")
```

## Save the objects

```{r}
save(crs.wintri, file = file.path(here::here(), "data/crs_wintri.rda"))
save(world, world.wintri, file = file.path(here::here(), "data/world.rda"))
buffer300 <- list(wintri = list(line = el300, polygon = e300, df = df300, crs = crs.wintri))
buffer20 <- list(wintri = list(line = el20, polygon = e20, df = df20, crs = crs.wintri))
save(buffer300, file = file.path(here::here(), "data/buffer300.rda"))
save(buffer20, file = file.path(here::here(), "data/buffer20.rda"))
```

