---
title: "4 - Offshore line parallel to coast"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{4 - Offshore line parallel to coast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example shows how to prepare a region that is parallel to the coastline. Once this is created, you can do geometric calculations in that area, like mean Chl-a.

Load the sample data.
```{r}
data("sample_raster", package="basics")
df <- sample_raster$df
ras <- sample_raster$raster
lons <- sample_raster$lons
lats <- sample_raster$lats
```

Load the needed packages for plotting.
```{r}
require(raster)
require(ggplot2)
```

## Download the world coastlines

For this buffer, I am using a fairly zoomed out version of the coastline. You could use the coastline from **rnaturalearth**.

```{r}
world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
world <- rgeos::gUnaryUnion(world)
```

```{r}
plot(raster::crop(world, raster::extent(lons[1], lons[2], lats[1], lats[2])))
```




## Add a buffer area

Here we add a buffer but it is hard to see at the world scale.
```{r}
buff1 <- raster::buffer(world, width = 1, dissolve = TRUE)
e<-erase(buff1, world)
plot(e, col="red", axes=TRUE)
```

Let's zoom in a bit.
```{r}
plot(raster::crop(e, raster::extent(lons[1], lons[2], lats[1], lats[2])), col="red", axes=TRUE)
```
One issue is that the units of the width are unclear. The help file for `buffer()` says that the units are meters when `+proj=longlat` as it is for `world`, but that is clearly not the case. The help file further implies that if the object is a `SpatialPolygon`, as it is in this case, the buffer is computed using **rgeos**. This would imply that `rgeos::gBuffer()` is used. But the `rgeos::gBuffer()` help file has no information on the units of `width`.  The distance appears to be in units of degree, and poking into the code for `.pointsBuffer()` ([here](https://rdrr.io/cran/raster/src/R/buffer.R) by searching on google), indicates that it sets `longlat = FALSE` if the data are lines or a polygon (not points). If `longlat=FALSE` then it uses units of degree. Units of degree are quite useless since this will not be area or distance uniform as we move around the map (north to south) or as our buffer rotates (from N-S to E-W).

How to do this? Looking at the `raster::buffer()` source code, it is clear that that function always uses units of degrees if the object is an **sp** object and not points. So let's try `rgeos::gBuffer()`. The help file is equally useless here regarding the units.

Here's the workflow:

`world` is in long-lat which is just x-y with longitude on x-axis and latitude on y-axis. Ok for plotting but in that x-y plane, 1-degree will have different kms in different places. We need to convert to a projection with meters. There are many we could use. Winkel tripel projection is a fairly standard one which is a compromise projection to preserve distance, area, and direction.

1. Define our projection string. `"+proj=wintri` means Winkel tripel projection. The `lon_0` and `lat_1` where it is centered. Let's keep at 0. `x_0` and `y_0` is the axis origin in the meters scale. We can ignore that too. `units` should be `m` or `km`. 
2. Change the `world` to this projection.
3. Calculate the buffer with that. Now it will use km as the unit.
4. Get rid of the world polygon from our buffer so our buffer is just the buffer.
5. Transform our buffer back to longlat so it is easier to plot and add to our raster plots.


```{r}
# Step 1
newcrs <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
# Step 2
x <- sp::spTransform(world, newcrs)
# Step 3
buff1 <- rgeos::gBuffer(x, width = 100)
e<-raster::erase(buff1, x)
```

It'll be helpful to have `lons` and `lats` in these new coordinates so we can crop.
```{r}
mpts <- SpatialPoints(cbind(lons, lats), proj4string=CRS("+proj=longlat"))
mpts <- sp::spTransform(mpts, newcrs)
```

Now we can plot as before:
```{r}
crop_e <- raster::crop(e, raster::extent(mpts))
plot(crop_e, col="red", axes=TRUE)
plot(x, add=TRUE, col="grey")
```

```{r}
plot(sp::spTransform(crop_e, crs(world)), axes=TRUE, col="red")
plot(world, add=TRUE, col="grey")
```

## Create a buffer that is a band offshore

Let's create a 100km wide buffer that is 250 to 350 km offshore. 250 km is about 3 degrees longitude offshore at 40 degrees latitude ([here is a calculator](https://www.nhc.noaa.gov/gccalc.shtml)).

```{r}
# Step 1
newcrs <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
# Step 2
x <- sp::spTransform(world, newcrs)
# Step 3
buff1 <- rgeos::gBuffer(x, width = 350, byid=TRUE)
buff2 <- rgeos::gBuffer(x, width = 250, byid=TRUE)
e<-raster::erase(buff1, buff2)
```

Now we can plot as before:
```{r}
plot(e, col="red", axes=TRUE, xlim=mpts@bbox[1,], ylim=mpts@bbox[2,])
plot(x, add=TRUE, col="grey")
```

## Add the SST raster to this

```{r}
plot(ras)
plot(sp::spTransform(crop_e, crs(world)), axes=TRUE, col="red", add=TRUE)
plot(world, add=TRUE, col="grey")
```

## Create a line around the world's coastlines at 300 km offshore

```{r}
# Step 1
newcrs <- "+proj=wintri +lon_0=0 +lat_1=0 +x_0=0 +y_0=0 +datum=WGS84 +units=km +no_defs"
# Step 2
x <- sp::spTransform(world, newcrs)
# Step 3
buff1 <- rgeos::gBuffer(x, width = 300, byid=TRUE)
e<-raster::erase(buff1, x)
```

If we wanted to get rid of the interior coastlines, we could use the `remove.holes()` function from **spatialEco**.
```{r}
nohole_e <- spatialEco::remove.holes(spatialEco::remove.holes(e))
```

Now we can plot as before and the holes that would be removed are the red lines.
```{r}
plot(e, border="red")
plot(nohole_e, add=TRUE)
plot(x, add=TRUE, col="grey")
```

```{r}
plot(e, border="red", axes=TRUE, xlim=mpts@bbox[1,], ylim=mpts@bbox[2,])
plot(x, add=TRUE, col="grey")
```
